---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# spaSim

<!-- badges: start -->
<!-- badges: end -->

The goal of spaSim is to facilicate tissue image simulations! It simulates cells
with 2D locations and cell types in a tissue. The available patterns include 
background cells, cell clusters, immune cell rings and vessels. It also enables 
simulations that generate a set of images in one run!

## Installation

You can install the development version of spaSim like so:

``` r
# install.packages("devtools")
devtools::install_github("fuerzhou/spaSim")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(spaSim)
set.seed(610)
mix_background <- TIS(n_cells = 5000, width = 2000, height = 2000, min_d = 10,
                      names_of_bg_cells = c("Tumour","Immune","Others"), 
                      proportions_of_bg_cells = c(0.1, 0.2, 0.7))
```

```{r spiat, include=FALSE}
library(SPIAT)
```

```{r}
plot_cell_categories(mix_background, categories_of_interest =c("Tumour","Immune"), 
                     colour_vector = c("red", "darkgreen"), feature_colname = "Phenotype")
```

